Inferring Galactic Parameters from Chemical Abundances with Simulation-Based Inference

### 2025-01-17 src/tex/ms.tex
[showyourwork] overleaf sync

\begin{figure}[]
     \centering
     \includegraphics[width=\columnwidth]{figures/ape_NN.png}
     \vspace{-.5cm}
     \caption{Absolute percentage error of the NN emulator for the \texttt{CHEMPY} simulator. The orange histogram shows the full distribution of percentage errors with the vertical dashed line indicating the median and the vertical dotted lines indicating the first and third quartile. The box plot on the top of the plot extends from the first quartile to the third quartile of the data, with a line at the median. The whiskers extend from the box to the farthest data point lying within $1.5\times$ the inter-quartile range from the box. The NN predicts abundances with an absolute percentage error far below typical observational errors.}
     \label{fig:ape_NN}
     \script{train_torch_chempy.py}
\end{figure}

@@

After a density estimator has been trained with simulated data to obtain a posterior, the estimator should be made subject to several diagnostic tests. This needs to be performed before being used for inference given the actual observed data. Posterior Predictive Checks provide one way to “critique” a trained estimator based on its predictive performance. Another important approach to such diagnostics is simulation-based calibration as developed by \citet{Cook2006} and \citet{Talts2018}.

@@

To perform SBC, we sample some $\theta_i^o$ values from the parameter prior of the problem at hand and simulate “observations” from these parameters: 
\begin{equation}
    x_i = \text{simulator}(\theta_i^o)
\end{equation}
Then we perform inference given each observation $x_i$ which produces a separate posterior $p_i(\theta|x_i)$ for each $x_i$. The key step for SBC is to generate a set of posterior samples $\{\theta\}_i$ from each posterior. We call this $\theta_i^s$, referring to $s$ samples from the posterior $p_i(\theta|x_i)$. Next, we rank the corresponding $\theta_i^o$ under this set of samples. A rank is computed by counting how many samples $\theta_i^s$ fall below their corresponding $\theta_i^o$ value \citep[see section 4.1 in][]{Talts2018}. These ranks are then used to perform the SBC check itself.

>>>>

\begin{figure}[]
     \centering
     \includegraphics[width=\columnwidth]{figures/ape_NN.pdf}
     \vspace{-.5cm}
     \caption{Absolute percentage error of the NN emulator for the \texttt{CHEMPY} simulator. The orange histogram shows the full distribution of percentage errors with the vertical dashed line indicating the median and the vertical dotted lines indicating the first and third quartile. The box plot on the top of the plot extends from the first quartile to the third quartile of the data, with a line at the median. The whiskers extend from the box to the farthest data point lying within $1.5\times$ the inter-quartile range from the box. The NN predicts abundances with an absolute percentage error far below typical observational errors.}
     \label{fig:ape_NN}
     \script{train_torch_chempy.py}
\end{figure}

@@

After a density estimator has been trained with simulated data to obtain a posterior, the estimator should be made subject to several diagnostic tests. This needs to be performed before being used for inference given the actual observed data. Posterior Predictive Checks provide one way to "critique" a trained estimator based on its predictive performance. Another important approach to such diagnostics is simulation-based calibration as developed by \citet{Cook2006} and \citet{Talts2018}.

@@

To perform SBC, we sample some $\theta_i^o$ values from the parameter prior of the problem at hand and simulate "observations" from these parameters: 
\begin{equation}
    x_i = \text{simulator}(\theta_i^o)
\end{equation}
Then we perform inference given each observation $x_i$ which produces a separate posterior $p_i(\theta|x_i)$ for each $x_i$. The key step for SBC is to generate a set of posterior samples $\{\theta\}_i$ from each posterior. We call this $\theta_i^s$, referring to $s$ samples from the posterior $p_i(\theta|x_i)$. Next, we rank the corresponding $\theta_i^o$ under this set of samples. A rank is computed by counting how many samples $\theta_i^s$ fall below their corresponding $\theta_i^o$ value \citep[see section 4.1 in][]{Talts2018}. These ranks are then used to perform the SBC check itself.
